Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-13 csf-2017/paper.tex
Added a couple of essential details to g1 lemma

\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f1: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[]$)
    pop $r_1$
    assert $r_1$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a stack-based calling convention.}
  \label{fig:prog-f1}
\end{figure}

\begin{figure}[htbp]
  \centering
\begin{lstlisting}
f2: malloc $r_l$ 1
    store $r_l$ 1
    fetch $r_1$ $\adv$
    call $r_1$($[]$,$[r_l]$)
    assert $r_l$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a calling convention not relyant
    on the stack}
  \label{fig:prog-f2}
\end{figure}

@@

\lau{in the following paragraph, we could make a reference to
  \textbf{invoking the adversary} from
  section~\ref{sec:stack-and-return-pointer} where the problem is
  first introduced. In fact, this paragraph is a reiteration of the
  problem in the aformentioned section.}
In \texttt{f3} there are two calls to an adversary and in order for
the assertion in the middle to succeed, they need to be
well-bracketed. If the adversary is able to store the return
pointer from the first call, so it is available for the second call, then it can be used
in a context where \texttt{f3} has $2$ on top of its stack which will cause the
assertion to fail. The return pointer is passed as a local
capability, so the only means the adversary has to store it is the
stack pointer. The stack pointer itself is a local capability,
so the adversary can only store it on the stack. This means that after
the adversary jumps to the return pointer, \texttt{f3} can
revoke it (by erasing it which happens as part of the
\texttt{scall}). In other words, the adversary cannot store the
return pointer for the second call, and it is possible to
prove that the assertion never fails.
\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}
\lau{Add something about the correctness lemma? Not sure if I have
  anything important to say.}
\begin{lemma}
  \label{lem:correctness-f3}
  Making similar assumptions about capabilities and linking as in
  Lemma~\ref{lem:correctness-f1}
  if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
  (\halted,m')$, then $m'(\flag) = 0$.
\end{lemma}

@@

\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \caption{A program that relies on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to $f4$.}
  \label{fig:prog-g1}
\end{figure}

@@

\begin{lemma}
  \label{lem:correctness-g1}
  Making similar assumptions about capabilities and linking as in
  Lemma~\ref{lem:correctness-f1} 
  if
  \[
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),
  \]
  then $m'(\flag) = 0$.
\end{lemma}

>>>>

\begin{figure}[t]
  \centering
  \begin{lstlisting}
f1: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[]$)
    pop $r_1$
    assert $r_1$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a stack-based calling convention.}
  \label{fig:prog-f1}
\end{figure}

\begin{figure}[t]
  \centering
\begin{lstlisting}
f2: malloc $r_l$ 1
    store $r_l$ 1
    fetch $r_1$ $\adv$
    call $r_1$($[]$,$[r_l]$)
    assert $r_l$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a calling convention not relyant
    on the stack}
  \label{fig:prog-f2}
\end{figure}

@@

\lau{in the following paragraph, we could make a reference to
  \textbf{invoking the adversary} from
  section~\ref{sec:stack-and-return-pointer} where the problem is
  first introduced. In fact, this paragraph is a reiteration of the
  problem in the aformentioned section.}
In \texttt{f3} there are two calls to an adversary and in order for
the assertion in the middle to succeed, they need to be
well-bracketed. If the adversary is able to store the return
pointer from the first call, so it is available for the second call, then it can be used
in a context where \texttt{f3} has $2$ on top of its stack which will cause the
assertion to fail. The return pointer is passed as a local
capability, so the only means the adversary has to store it is the
stack pointer. The stack pointer itself is a local capability,
so the adversary can only store it on the stack. This means that after
the adversary jumps to the return pointer, \texttt{f3} can
revoke it (by erasing it which happens as part of the
\texttt{scall}). In other words, the adversary cannot store the
return pointer for the second call, and it is possible to
prove that the assertion never fails. The correctness lemma for \texttt{f3} is similar to the two previous lemmas:
\begin{figure}[t]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

\begin{lemma}
  \label{lem:correctness-f3}
  Making similar assumptions about capabilities and linking as in
  Lemma~\ref{lem:correctness-f1}
  if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
  (\halted,m')$, then $m'(\flag) = 0$.
\end{lemma}

@@

\begin{figure}[t]
  \centering
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \caption{A program that relies on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to $f4$.}
  \label{fig:prog-g1}
\end{figure}

@@

\begin{lemma}
  \label{lem:correctness-g1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
    \end{array}
\]
  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1} 
  if
  \[
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),
  \]
  then $m'(\flag) = 0$.
\end{lemma}
